BIP32-Ed25519++: A More Secure Hierarchical Deterministic Wallet Standard for Ed25519 Keys

### 2024-06-13 paper.tex
minor wording/clarity changes

In the basic case, a single seed phrase corresponds to a single address, requiring users to keep track of many seed phrases should they wish to use separate addresses for different application domains: making and receiving payments, holding NFTs, participating in a DAO, DeFi, gaming, etc.

@@

Bitcoin relies on the Secp256k1~\cite{Secp256k1} elliptic curve, which has a linear key space.
A new public key can be formed by adding $[x]B$ to the public key $[k]B$, where $x$ and $k$ are private keys (scalars), and $B$ is the base point.
This results in a new public key whose private key is $[x] + [k] \bmod q$, where $q$ is the order of the field.

@@

Unfortunately, private keys in Ed25519 have to be derived by hashing a private ``seed'' and manipulating the bits.
RFC\-8032~\cite{RFC8032} specifies that in the 32-byte private key:
\begin{itemize}[label=\textendash, itemsep=-0.5em]
\item Bits 0, 1, 2 are set to 0, ensuring that the key is a multiple of 8, thus avoiding small-subgroup attacks.
\item Bit 255 is set to 0, to ensure that the key is less than the order of the elliptic-curve group.
\item Bit 254 is set to 1, to ensure that the Montgomery ladder computation used in scalar multiplication runs in constant time.
\end{itemize}
So, by definition, the key space for Ed25519 is non-linear relative to the group order.

@@

In this paper an improvement to their method is proposed that provides a more secure way to derive Ed25519 keys, at the cost of reducing the number of levels of child keys.

@@

While reviewing the BIP32-Ed25519 paper~\cite{BIP32-Ed25519}, two issues were identified, the first minor and the other major.

@@

While deriving a new private or public child key, it is possible to produce a child that is unsafe to use, in the following two cases:
\begin{itemize}[label=\textendash, itemsep=-0.5em]
\item $k_L$ is divisible by the order of the base point,
\item $A_i$ is the identity point.
\end{itemize}

@@

As was mentioned in \cref{subsec:elliptic_curves_and_their_key_spaces}, for Ed25519 private-keys scalars, one must ensure that bits 0, 1, 2, 3 and 255 are 0, and that bit 254 is 1.
The paper's method for ensuring this is to produce the child private key and public key as follows.

@@

To understand how many levels can be safely performed, consider that the root key~$a$ can be expressed as $a = 2^{254} + 8b < 2^{254} + 2^{253}$, where $b < 2^{250}$ due to the bit manipulation and the way it is generated.

@@

The purpose of our modifications is to make derived keys more randomized and less related, so that there is less security degradation if some private keys are compromised.

@@

Similarly, $g=16$ is the next value that breaks at the byte boundary (30 bytes) and provides more than enough levels.
However, it randomizes only 240 out of the 251 bits available in an Ed25519 scalar.

@@

It also becomes apparent that the quoted maximum depth~$2^{20}$ from~\cite{BIP32-Ed25519} is smaller than the actual number of safe levels, which is at least $2^{26}$.

@@

In particular, setting $g=9$ allows for the most secure implementation that also satisfies the requirements of the most common derivation schemes.

>>>>

In the basic case, a single seed phrase corresponds to a single address, requiring users to keep track of many seed phrases should they wish to use separate addresses for different application domains: e.g., making and receiving payments, holding NFTs, participating in a DAO, DeFi, gaming, etc.

@@

Bitcoin relies on the Secp256k1~\cite{Secp256k1} elliptic curve, which has a linear key space.
A new public key can be formed by adding $[x]B$ to the public key $[k]B$, where $x$ and $k$ are private keys (scalars), and $B$ is the base point.
This results in a new public key whose private key is $[x] + [k] \bmod q$, where $q$ is the order of the scalar field.

@@

However, private keys in Ed25519 have to be derived by hashing a private ``seed'' and manipulating the bits.
RFC\-8032~\cite{RFC8032} specifies that in the 32-byte private key:
\begin{itemize}[label=\textendash, itemsep=-0.5em]
\item Bits 0, 1, 2 are set to 0, ensuring that the key is a multiple of 8, thus avoiding small-subgroup attacks.
\item Bit 255 is set to 0, to ensure that the key is less than the order of the elliptic-curve group.
\item Bit 254 is set to 1, to ensure that the Montgomery ladder computation used in scalar multiplication runs in constant time.
\end{itemize}
So, by definition, the key space for Ed25519 is non-linear relative to the group order: the sum of two private keys (modulo the group order) is not necessarily a valid private key.

@@

This paper proposes an improvement to the method from~\cite{BIP32-Ed25519} that provides a more secure way to derive Ed25519 keys, at the cost of reducing the number of levels in the hierarchy of child keys.

@@

While reviewing the BIP32-Ed25519 paper~\cite{BIP32-Ed25519}, two issues were identified, the first minor and the other significant.

@@

When deriving a new private or public child key, it is possible to produce a child that is unsafe to use, in the following two cases:
\begin{itemize}[label=\textendash, itemsep=-0.5em]
\item $k_L$ is divisible by the order of the base point,
\item $A_i$ is the identity point.
\end{itemize}

@@

As mentioned in \cref{subsec:elliptic_curves_and_their_key_spaces}, for Ed25519 private-keys scalars, one must ensure that bits 0, 1, 2, 3 and 255 are 0, and that bit 254 is 1.
The method proposed in~\cite{BIP32-Ed25519} for ensuring this is to produce the child private key and public key as follows.

@@

To understand how many levels can be safely derived, consider that the root key~$a$ can be expressed as $a = 2^{254} + 8b < 2^{254} + 2^{253}$, where $b < 2^{250}$ due to the bit manipulation and the way it is generated.

@@

The purpose of our modifications is to add more randomization to child keys, so that there is less security degradation if some private keys are compromised.

@@

Similarly, $g=16$ is the next value that breaks at the byte boundary (30 bytes), and it provides more than enough levels.
However, it randomizes only 240 out of the 251 bits available in an Ed25519 scalar.

@@

From this more general setup it is also apparent that the quoted maximum depth~$2^{20}$ from~\cite{BIP32-Ed25519} is smaller than the actual number of safe levels, which is at least $2^{26}$.

@@

In particular, setting $g=9$ allows for the most secure implementation that also satisfies the requirements of Ed25519 and the most common derivation schemes.
